Managing variability in software systems

### 2021-05-02 Paper.tex
add discussion

\section{Variability modeling (1 page) JR}
As mentioned there are two main approaches for managing variability. For each of which we will discuss the advantages and disadvantages and create a broad overview over the different modeling approaches.\\\\
\textbf{Feature modeling}\\
Features in the context of feature modeling are behaviours, characteristics or functionalities which are observable by the user of the software system. They can be used to define commonalities and differences between systems and their variants, as well as determine mandatory, optional or alternative aspects of them\cite{b8}. A broadly used approach of displaying features and their relations is hierarchically structured diagram which is called a features diagrams.\\\\
A big advantage of feature models, especially in the shape of feature diagrams, is a possible communication enhancement between stakeholders. This can be achieved by providing a overview of all functionalities which are offered to the user by the system. Due to the nature of the model regarding structure and due to the fact that it displays if and which planned features are already provided by another system or system family, this may improve the cost and time estimation for developing the software system. \\\\
The essence of feature modeling is the splitting of features into distinct objects which have a certain relation to each other. There are several types of features mentioned in \cite{b8}, named \textit{mandatory, optional and alternative features}. Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models.\cite{b9}\\\\ 
If a feature is mandatory it has to be selected if its parent feature is selected, the type of the parent feature does not matter. A optional feature can be selected freely but there may be constraints from composition rules (e.g. a feature can only be choosen as part of a feature sets). Alternative features are can be pictured as a choice box, so one can choose \textit{one} item from a feature set.\\\\
This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features. \cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.\\\\
In Fig. 1, one can see a small feature diagram of a car, it consists of one root node, that describes the domain we are modeling, in this case the car. The previous elaborated mandatory features are the engine and seats. Furthermore there is the option to install a cruise control and the option for a turbocharger these features are completely independet and optional. Last but not least one can choose between normal leather seats or alcantara ones but not both. Which this simplified model of a car we have all three types of features modeled

\textbf{Decision modeling}\\
asdfasdf

@@

\section{Discussion/Challenges}
Discuss why the variability of systems is hard to tackle and why it is not always successful.
(1/4 page) JR

>>>>

\section{Variability modeling (1 page) JR}
As mentioned there are two main approaches for managing variability. For each of which we will discuss the advantages and disadvantages and create a broad overview over the different modeling approaches.\\\\
\textbf{Feature modeling}\\
Features in the context of feature modeling are behaviours, characteristics or functionalities which are observable by the user of the software system. They can be used to define commonalities and differences between systems and their variants, as well as determine mandatory, optional or alternative aspects of them\cite{b8}. A broadly used approach of displaying features and their relations is hierarchically structured diagram which is called a features diagrams.\\\\
A big advantage of feature models, especially in the shape of feature diagrams, is a possible communication enhancement between stakeholders. This can be achieved by providing a overview of all functionalities which are offered to the user by the system. Due to the nature of the model regarding structure and due to the fact that it displays if and which planned features are already provided by another system or system family, this may improve the cost and time estimation for developing the software system. \\\\
The essence of feature modeling is the splitting of features into distinct objects which have a certain relation to each other. There are several types of features mentioned in \cite{b8}, named \textit{mandatory, optional and alternative features}. Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models.\cite{b9}\\\\ 
If a feature is mandatory it has to be selected if its parent feature is selected, the type of the parent feature does not matter. A optional feature can be selected freely but there may be constraints from composition rules (e.g. a feature can only be choosen as part of a feature sets). Alternative features are can be pictured as a choice box, so one can choose \textit{one} item from a feature set.\\\\
This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features.\cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.\\\\
In Fig. 1, one can see a small feature diagram of a car, it consists of one root node, that describes the domain we are modeling, in this case the car. The previous elaborated mandatory features are the engine and seats. Furthermore there is the option to install a cruise control and the option for a turbocharger these features are completely independet and optional. Last but not least one can choose between normal leather seats or alcantara ones but not both. Which this simplified model of a car we have all three types of features modeled.\\\\
\textbf{Decision modeling}\\
In decision modeling the variability is described with decision instead of features. A variable feature in feature modeling is equal to a decision and vice versa. Contrary to feature modeling decision modeling is unable to describe the commonalities of systems.\\\\
Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependicies. In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy. \cite{b6}\\\\
Same as feature modeling, decision modeling has been expanded over time to enable the application coverage of DM be almost the same as FM.\cite{b6}\\\\

@@

\section{Discussion/Challenges (1/4 page) JR}
The research field of variability had a increasing trend of published paper per year from 2000 until 2008. After that it stabilized between 25-30 new research paper per year from 2008 until 2011. One explanation of this rise in interest in the scientific community could be that the usage of variability was becoming more broadly in dynamic systems. \cite{b4}\\\\
Another aspect is that feature modeling is in its core a simple concept. Even though its impact in the industry is pretty big considering it is the most popular concept for variability modeling in software systesm until today.\cite{b9} One can say that because it is such a simple concept and due to the big amount of research in this field we understand the subject pretty well by now.\\\\
The currently existing methods also have their limitations. It is difficult to guarantee correctness and consistency, authors have to urge for verification and consistency checking. Performance overheads, runtime upgradability and the lack of support regarding security needs are some areas where the described methods are limited. Sometimes the handling of variability can create poor user-friendliness and therefore decrease usebility of a system. \cite{b4}\\\\
There are ways to decrease the negative impact of these aspects e.g. correctness and consistency can be checked by various methods such as modeling guidelines customized for specific needs, together with a reviewing process where changes to the model are checked by another person to ensure the quality of the metioned parameters.
